---
title: "Tabulating Sociodemographic Data for Final Report to NIJ"
date: "Created: 2019-01-25 <br> Updated: `r Sys.Date()`"
output: 
  html_notebook:
    toc: true
    toc_float: true
    css: custom-css.css
---

# Overview

Below we tabulate select sociodemographic variables for inclusion our summary report to NIJ.


# Load packages and data

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
Sys.setenv(TZ = "US/Central")
```

```{r message=FALSE}
library(tidyverse)
library(bfuncs)
```

medstar_epcr_02_variable_management.feather was created in data_medstar_epcr_02_variable_management.Rmd

```{r}
medstar_epcr <- feather::read_feather(
  "/Volumes/sph_research/Detect/one_year_data/medstar_epcr_02_variable_management.feather"
)
```

```{r}
about_data(medstar_epcr) # 28,228 observations and 56 variables
```

[top](#top)










# Double-check dates

Check earliest and latest dates in the data

```{r}
medstar_epcr %>% 
  filter(screened == 1) %>% 
  summarise(
    `Earliest Screened` = min(arrival_time), # 2017-02-01 01:54:34
    `Latest Screened`   = max(arrival_time)  # 2018-02-28 23:49:52
  )
```

[top](#top)










# Screenings

Count the number of screenings conducted

```{r}
medstar_epcr <- medstar_epcr %>% 
  filter(screened == 1)
```

```{r}
medstar_epcr %>% 
  summarise(
    `DETECT Screenings` = format(n(), big.mark = ",") # 24,007
  )
```

[top](#top)










# Unique people

Reduce the data to one row per person -- defined by unique values of name and date of birth.

```{r}
medstar_epcr <- medstar_epcr %>% 
  mutate(person = paste(name_first, name_last, birth_year, birth_mnth, birth_day, sep = "_")) %>% 
  group_by(person) %>% 
  filter(row_number() == 1) %>% 
  ungroup()
```

```{r}
about_data(medstar_epcr) # 16,261 observations and 57 variables
```

[top](#top)










# Calculate statistics of interest

## Age

```{r}
medstar_epcr %>% 
  mean_table(age) %>% 
  format_table(stats = "n and mean")
```

## Gender

```{r}
medstar_epcr %>% 
  group_by(gender) %>% 
  freq_table() %>% 
  format_table(stats = "n and percent")
```

## Race and Ethnicity

```{r}
medstar_epcr %>% 
  group_by(race_8cat) %>% 
  freq_table() %>% 
  format_table(stats = "n and percent")
```

[top](#top)










# Session information

```{r echo=FALSE}
rm(list = ls())
```

```{r echo=FALSE}
sessionInfo()
```