---
title: "Estimates for the Follow-up Flow Chart"
date: "Created: 2019-02-17 <br> Updated: `r Sys.Date()`"
output: 
  html_notebook:
    toc: true
    toc_float: true
    css: custom-css.css
---

# Overview

I'm working on the flow chart for the follow-up interview process. I need to estimate a couple things like the number of sreenings MedStar will conduct each week.


# Load packages and data

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
Sys.setenv(TZ = "US/Central")
```

```{r message=FALSE}
library(tidyverse)
library(bfuncs)
```

medstar_epcr_02_variable_management.feather was created in data_medstar_epcr_02_variable_management.Rmd

```{r}
medstar <- feather::read_feather("/Volumes/DETECT/one_year_data/medstar_epcr_02_variable_management.feather")
```

```{r}
about_data(medstar) # 28,228 observations and 56 variables
```

[top](#top)










# DETECT screenings

How many DETECT screenings do we anticipate doing each week?

Our best guess should come from last year's data.

```{r}
medstar %>% 
  filter(screened) %>% 
  select(arrival_time) %>% 
  mutate(
    year = lubridate::year(arrival_time),
    week = lubridate::week(arrival_time),
    year_week = paste(year, week, sep = "_")
  ) %>% 
  group_by(year_week) %>% 
  summarise(n = n()) %>% 
  summarise(`Mean Screenings Per Week` = mean(n) %>% round())
```

[top](#top)










# Session information

```{r echo=FALSE}
rm(list = ls())
```

```{r echo=FALSE}
sessionInfo()
```