---
title: "Manage Variables in APS Data"
date: "Created: 2019-01-25 <br> Updated: `r Sys.Date()`"
output: 
  html_notebook:
    toc: true
    toc_float: true
    css: custom-css.css
---

# Overview

* The raw APS data was imported in data_aps_01_import.Rmd

* In this file we prepare the data for analysis


# Load packages and data

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
Sys.setenv(TZ = "US/Central")
```

```{r message=FALSE}
library(tidyverse)
library(bfuncs)
```

aps_01_import.feather was created in data_aps_01_import.Rmd

```{r}
aps <- feather::read_feather("/Volumes/SPH/Research/Detect/one_year_data/aps_01_import.feather")
```

```{r}
about_data(aps) # 18,080 observations and 47 variables
```

[top](#top)










# Standardize character strings

Because we will merge this data with other data sets in the future based on character strings (e.g., name), we need to go ahead and standardize their formats here. This will prevent mismatches during the merges. Specifically, we:

1. Transform all characters to lower case   
2. Remove trailing spaces (e.g., "John Smith ")   
3. Remove double spaces (e.g., "John  Smith")

```{r}
aps <- aps %>% 
  mutate_if(is.character, tolower) %>% 
  mutate_if(is.character, stringr::str_replace, "[[:blank:]]$", "") %>%
  mutate_if(is.character, stringr::str_replace_all, "[[:blank:]]{2,}", " ")
```

4. Remove any special characters (e.g., hyphens, periods)   

```{r}
vars <- quos(name_last, name_first, name_full)

aps <- aps %>% 
  mutate_at(vars(!!! vars), stringr::str_replace_all, "[^a-zA-Z\\d\\s]", " ")

rm(vars)
```

```{r}
about_data(aps) # 18,080 observations and 47 variables
```

[top](#top)










# Clean city

View current values for city

```{r eval=FALSE}
aps %>%
  group_by(city) %>%
  freq_table()
```

Need to move city value to street. Need to do so without using identifiers in the code. Therefore, adding a temporary row variable.

```{r}
aps <- aps %>% 
  mutate(row = row_number()) %>% 
  select(row, everything())
```

```{r}
aps <- aps %>% 
  mutate(
    street = if_else(row == 9868, city, street),
    city = if_else(row == 9868, NA_character_, city)
  )
```

Correct misspellings

```{r}
aps <- aps %>% 
  mutate(
    city = case_when(
      city == "cleburn" ~ "cleburne",
      city == "dalworthington garde" ~ "dalworthington garden",
      city == "ft worth" ~ "fort worth",
      city == "n richland hills" ~ "north richland hills",
      city == "unknown" ~ NA_character_,
      city == "wht settlemt" ~ "white settlement",
      TRUE ~ city
    )
  )
```

Drop row

```{r}
aps <- select(aps, -row)
```

```{r}
about_data(aps) # 18,080 observations and 47 variables
```

[top](#top)










# Separate dob's and street addresses

* We also separate dob into its component parts: month, day, year.

* We also separate the street address into the number part and the street name part

```{r}
aps <- aps %>%
  mutate(
    birth_mnth          = lubridate::month(dob),
    birth_day           = lubridate::day(dob),
    birth_year          = lubridate::year(dob),
    address_num         = stringr::str_extract(street, "^\\d{1,5}") %>% as.numeric(),
    address_street_name = stringr::str_trim(str_replace(street, "^\\d{1,5}", ""))
  )
```

```{r}
about_data(aps) # 18,080 observations and 52 variables
```

[top](#top)










# MedStar service area

According to [MedStar's website](http://www.medstar911.org/service-area), the cities they serve include: Blue Mound, Burleson, Edgecliff Village, Forest Hill, Fort Worth, Haltom City, Haslet, Lakeside, Lake Worth, River Oaks, Saginaw, Sansom Park, Westover Hills, Westworth Village, White Settlement.

What cities appear in the data?

```{r eval=FALSE}
aps %>% 
  group_by(city) %>% 
  freq_table()
```

There are a lot of cities included in this data that our outside of MedStar's service area. Create a dummy variable to indicate cities included in MedStar's coverage area.



```{r}
aps <- aps %>% 
  mutate(medstar_service_area = if_else(city %in% c(
    "blue mound", "burleson", "edgecliff village", "forest hill", "fort worth", 
    "haltom city", "haslet", "lake worth", "lakeside", "river oaks", "saginaw", 
    "sansom park", "westover hills", "westworth village", "white settlement"
    ), 1, 0, NA_real_
  ))
```

Note: There are no APS investigations with a city address of Edgecliff Village or Westover Hills.

```{r}
about_data(aps) # 18,080 observations and 53 variables
```

Add a little information about the area this data covers

What counties are represented in this data?

```{r echo=FALSE}
aps %>% 
  mutate(
    `Medstar Service Area` = if_else(
      medstar_service_area == 0,
      "Outside MedStar's Service Area",
      "Inside MedStar's Service Area",
      NA_character_
    )
  ) %>% 
  group_by(`Medstar Service Area`) %>% 
  summarise(
    Counties = unique(county) %>% sort() %>% paste(collapse = ", ")
  )
```


How many cities are inside and outside of the MedStar Service Area?

```{r echo=FALSE}
aps %>% 
  mutate(
    `Medstar Service Area` = if_else(
      medstar_service_area == 0,
      "Outside MedStar's Service Area",
      "Inside MedStar's Service Area",
      NA_character_
    )
  ) %>% 
  group_by(`Medstar Service Area`) %>% 
  summarise(
    `Number of Cities` = unique(city) %>% length()
  )
```

[top](#top)










# Create city_fw

```{r}
aps <- aps %>% 
  mutate(city_fw = if_else(city == "fort worth", 1, 0))
```

```{r}
about_data(aps) # 8,927 observations and 62 variables
```

[top](#top)










# Create DETECT identifier

Create a dummy variable corresponding to the dates when MedStar was using the DETECT tool.

```{r}
aps <- aps %>% 
  mutate(
    detect = case_when(
      as.Date(intake_date) >= "2015-09-17" & as.Date(intake_date) <= "2015-10-26" ~ 1,
      as.Date(intake_date) >= "2017-02-01"                                        ~ 1,
      TRUE                                                                        ~ 0
    )
  )
```

```{r}
about_data(aps) # 18,080 observations and 55 variables
```

[top](#top)










# Create study week variable

## Data time period

What time period does the APS data cover?

```{r echo=FALSE}
aps %>% 
  summarise(
    `Earliest Date` = min(intake_date),
    `Latest Date` = max(intake_date)
  )
```

## Create a week number variable.

```{r}
aps <- aps %>% 
  mutate(
    intake_year = format(intake_date, "%y"),
    year_add = case_when(
      intake_year == "14" ~ (0 * 52),
      intake_year == "15" ~ (1 * 52),
      intake_year == "16" ~ (2 * 52),
      intake_year == "17" ~ (3 * 52),
      intake_year == "18" ~ (4 * 52)
    ),
    week = lubridate::week(intake_date),
    study_week = week + year_add,
    study_week = study_week - 52 # To make it base 1
  ) %>% 
  arrange(intake_date)
```

```{r}
# Data checks
# aps_its %>% select(intake_date, intake_year, year_add, week, study_week, detect)
```

During which study weeks did DETECT screening occur?

```{r echo=FALSE}
aps %>% 
  filter(detect == 1) %>% 
  # For data checking
  # select(intake_date, study_week)
  pull(study_week) %>% 
  unique()
```

Drop unneeded variables

```{r}
aps <- aps %>% 
  select(-year_add, -week)
```

```{r}
about_data(aps) # 18,080 observations and 57 variables
```

[top](#top)










# Create dummy variable for EMS reporter

```{r}
aps <- aps %>% 
  mutate(ems = if_else(reporter == "health care providers/staff -- ems/emt", 1, 0))
```

```{r}
about_data(aps) # 18,080 observations and 58 variables
```

[top](#top)










# Valid/Invalid summary

Create a new variable that summarizes the results of APS investigations

It is possible for there to be more than one allegation associated with each APS investigation.

We want a single outcome per investigation/pcr combination in order to create a confusion matrix for each screening item.

To do that, we will create a new variable called "any_valid" that is equal to 1 if the disposition of _any_ allegation was valid, and 0 if the disposition of _all_ allegations were invalid.   
* If even one disposition is valid, then any valid equals 1   
* If all dispositions are invalid, then any valid equals 0   
* If all dispositions are other, then any valid equals NA
* If one disposition is other, and the rest are invalid, then any valid equals NA   

```{r}
aps <- aps %>% 
  mutate(
    total_valid   = select(., starts_with("valid")) %>% rowSums(),
    total_invalid = select(., starts_with("invalid")) %>% rowSums(),
    any_valid     = as.numeric(total_valid > 0),
    all_invalid   = as.numeric(total_valid == 0 & total_invalid > 0)
  )
```

Create 3-level summary variable

1. Any valid finding    
2. All invalid findings    
3. Unable to determine or other finding

```{r}
aps <- aps %>% 
  mutate(
    valid_3cat = case_when(
      any_valid == 1 ~ 1,
      all_invalid == 1 ~ 2,
      TRUE ~ 3
    )
  )
```

```{r}
# Data checks
# check <- aps %>% 
#   select(closure_reason:total_other_dispositions, total_valid:valid_3cat)
```

```{r}
about_data(aps) # 18,080 observations and 63 variables
```

[top](#top)










# Save data

```{r}
feather::write_feather(aps, "/Volumes/sph_research/Detect/one_year_data/aps_02_variable_management.feather")
```

[top](#top)










# Session information

```{r echo=FALSE}
rm(list = ls())
```

```{r echo=FALSE}
sessionInfo()
```