---
title: "Figure: Change in Reporting Over Time"
date: "Created: 2019-01-29 <br> Updated: `r Sys.Date()`"
output: 
  html_notebook:
    toc: true
    toc_float: true
    css: custom-css.css
---

# Overview

In this file we visualize the changes in reporting to APS over time for EMS and non-EMS reporters.


# Load packages and data

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
Sys.setenv(TZ = "US/Central")
```

```{r message=FALSE}
library(tidyverse)
library(gridExtra)
```

```{r}
source("../r/theme_bfuncs.r")
```

aps_ddd.feather was created in data_aps_03_process_for_ddd.Rmd.

```{bash}
open 'smb://uctnascifs.uthouston.edu/sph_research/DETECT'
```

```{r}
aps_ddd <- feather::read_feather("/Volumes/DETECT/one_year_data/aps_ddd.feather")
```

```{r}
dim(aps_ddd) # 118,080    45
```

[top](#top)


# Data management

* Looks like we have incomplete informtion for week 165. Let's drop it from the ITS analysis.

* Create groups to plot by

```{r}
aps_ddd_w_groups <- aps_ddd %>% 
  # Looks like we have incomplete informtion for week 165
  filter(study_week < 165) %>% 
  # Create groups
  mutate(
    group = case_when(
      ems == 0 & medstar_service_area == 0 ~ "Non-medics, Outside MedStar Service Area",
      ems == 1 & medstar_service_area == 0 ~ "Medics, Outside MedStar Service Area",
      ems == 0 & medstar_service_area == 1 ~ "Non-medics, Inside MedStar Service Area",
      ems == 1 & medstar_service_area == 1 ~ "Medics, Inside MedStar Service Area"
    ) %>% 
      as.factor()
  )
```


## Calculate number of reports by group

```{r}
weekly_reports_by_group <- aps_ddd_w_groups %>%
  group_by(study_week, group) %>% 
  summarise(reports = n()) %>% 
  ungroup()
```


## Add zero report weeks

Any combinations of study_week and group that didn't occur in the data are non-existant. Really we want them to be 0.

Add week/group combinations that had zero reports

```{r}
weekly_reports_by_group <- data.frame(
    # Create data frame with all weeks and groups
    study_week = seq(
      min(aps_ddd_w_groups$study_week), 
      max(aps_ddd_w_groups$study_week)
    ) %>% 
      rep(each = 4),
    group = levels(aps_ddd_w_groups$group),
    reports = 0
  ) %>% 
  
  # Which week/group combinations are not in the data?
  anti_join(weekly_reports_by_group, by = c("study_week", "group")) %>% 
  
  # Merge those rows into the data
  full_join(weekly_reports_by_group, by = c("study_week", "group", "reports")) %>% 
  
  # Arrange the data
  arrange(study_week, group)
```


## Add detect study periods

Also add mean reports in each period

```{r}
weekly_reports_by_group <- weekly_reports_by_group %>% 
  mutate(
    period = case_when(
      between(study_week, 1,  37)  ~ 1, # Pre-DETECT
      between(study_week, 38, 43)  ~ 2, # DETECT 5-week
      between(study_week, 44, 108) ~ 3, # Washout
      study_week > 108             ~ 4  # DETECT 1-year
    )
  ) %>% 
  group_by(group, period) %>%
  mutate(mean_reports = mean(reports)) %>% 
  ungroup()
```

[top](#top)










# Create plots

It isn't currently possible to plot different y-axis scales for the first and second row of the faceted plot using ggplot2 alone. Therefore, we have to create row 1 and row 2 of the plot separately and then combine them using grid extra.

```{r}
# Separate the data
medic_weekly_reports_by_service_area <- weekly_reports_by_group %>% 
  filter(stringr::str_detect(group, "Medics"))

non_medic_weekly_reports_by_service_area <- weekly_reports_by_group %>% 
  filter(stringr::str_detect(group, "Non-medics"))
```

```{r}
# Create the plot function
plot_weekly_reports <- function(.data) {
  .data %>% 
    ggplot(aes(study_week, reports)) +
      
      # Add DETECT study periods
      geom_rect(aes(xmin = 38, xmax = 43, ymin = -Inf, ymax = Inf), fill = "lightyellow") +
      geom_rect(aes(xmin = 109, xmax = 164, ymin = -Inf, ymax = Inf), fill = "lightyellow") +
      
      # Add reports by week
      geom_point(size = 0.2) +
      geom_line(group = 1, color = "gray70") +
    
      # Add mean lines
      geom_line(aes(y = mean_reports, group = period), color = "red", size = 2) +
    
      # Add facets
      facet_wrap(vars(group), scales = "fixed") +
    
      # Adjust theme
      theme_bfuncs() +
      theme(
        panel.border = element_rect(fill = NA, colour = "black"),
        strip.text = element_text(size = rel(1.))
      ) +
      
      # Adjust labels
      labs(
        x = "Study Week", 
        y = "Number of Reports"
      )
}
```

```{r}
# Create plots
medic_plot <- plot_weekly_reports(medic_weekly_reports_by_service_area)
non_medic_plot <- plot_weekly_reports(non_medic_weekly_reports_by_service_area)
```

```{r}
# Combine the two plots
fig_aps_ddd_change_in_reporting <- gridExtra::grid.arrange(medic_plot, non_medic_plot)
```

```{r}
ggsave(
  filename = "../images/fig_aps_ddd_change_in_reporting.png",
  plot     = fig_aps_ddd_change_in_reporting,
  device   = "png",
  width    = 11.43
)
```

2020-02-21: Create an eps version of the file for the JAMA R&R

```{r}
ggsave(
  filename = "../images/fig_aps_ddd_change_in_reporting.eps",
  plot     = fig_aps_ddd_change_in_reporting,
  device   = "eps",
  width    = 11.43
)
```

[top](#top)


# Mean age

```{r}
bfuncs::mean_table(aps_ddd, age)
```











# Session information

```{r echo=FALSE}
rm(list = ls())
```

```{r echo=FALSE}
sessionInfo()
```