---
title: "Creating Patient Symptom Table"
date: "Created: 2019-01-31 <br> Updated: `r Sys.Date()`"
output: 
  html_notebook:
    toc: true
    toc_float: true
    css: custom-css.css
---

# Overview

Details on the construction of the MedStar EPCR Patient symptom table.   126 unique symptoms were recorded for 2853 calls.  A total of 4086 symptoms were recorded among the calls.


# Load packages and data

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
Sys.setenv(TZ = "US/Central")
```

```{r message=FALSE}
library(tidyverse)
library(bfuncs)
```

```{r eval=FALSE}
symptoms <- feather::read_feather("/Users/jaredwiegand/Desktop/DETECT Repository/symptoms.feather")
```

* Adding file path for DETECT data server.    
* Brad emailed Chris Harvey on 2019-02-11 to make Jared an authorized user.

```{r}
symptoms <- feather::read_feather("/Volumes/DETECT/one_year_data/symptoms.feather")
```

```{r}
about_data(symptoms) # 28,228 observations and 2 variables
```

[top](#top)










# Simplify symptom list

Currently, the symptom list includes upper and lower case letters, spaces, and special characters. These are not characteristics we want on our variable names, and they make matching with regular expressions more difficult. We want to do both of those things below. So, here we will:

* Convert all letters to lower case   
* Replace spaces with underscores   
* Replace parentheses with underscores   
* Replace periods with underscores  
* Replace commas with underscores  
* Replace all instances of 2 or more consecutive underscores we created in the process with single underscores   
* Remove trailing underscores   

```{r}
simplify_symptom_strings <- function(x) {
  x %>% 
    tolower() %>% 
    stringr::str_replace_all(pattern = "\\s", replacement = "_") %>% # white space
    stringr::str_replace_all(pattern = "[\\(|\\)]", replacement = "_") %>% # parentheses
    stringr::str_replace_all(pattern = "\\.", replacement = "_") %>% # periods
    stringr::str_replace_all(pattern = "\\,(?=\\_)", replacement = "_") %>% # commas - not separating symptoms
    stringr::str_replace_all(pattern = "\\_{2,}", replacement = "_") %>% # multiple underscores
    stringr::str_replace_all(pattern = "\\_(?=\\')", replacement = "") # trailing underscores
}
```

```{r}
symptoms <- symptoms %>% 
  mutate(symptoms_simplified = simplify_symptom_strings(symptoms))
```

```{r}
about_data(symptoms) # 28,228 observations and 3 variables
```

[top](#top)










# Find all symptoms

* Figure out how to break character stings into component symptoms   
* Get list of all unique symptoms   
* Turn symptoms into binary variables   
* Create new table of symptoms   
* Create two-way table of symptoms by incident complaint   


## Create List of Unique Symptoms

```{r}
symptom_list <- symptoms %>% 
  filter(!is.na(symptoms_simplified)) %>% # Remove NA
  pull(symptoms_simplified) %>% # Coerce to vector 
  unique() %>% # Gets unique groups of symptoms
  stringr::str_split(pattern = "'',''") %>% # Splits character string at '',''
  unlist() %>% # Coerce from list to vector
  stringr::str_remove_all("'") %>% # Remove remaining single quotes
  unique() # Get unique symptoms
```


## Create dummy variables for each symptom, by the same name

```{r}
for (i in seq_along(symptom_list)) {
  symptoms[symptom_list[i]] <- NA
}
rm(i)
```

```{r}
about_data(symptoms) # 28,228 observations and 129 variables
```


## Assign values for each symptom dummy variable

```{r}
for (i in seq_along(symptom_list)) {
  # For each variable in symptoms that shares the name of the i'th value in symptom_list
  symptoms[symptom_list[i]] <- if_else( 
    # If the name of the variable is found anywhere in symptoms_simplified
    # Set the value of that variable to 1 in that row
    # Otherwise set it to 0
    stringr::str_detect(
      string = symptoms$symptoms_simplified, 
      pattern = symptom_list[i]
    ), 1, 0
  )
}
rm(i)
```

```{r}
about_data(symptoms) # 28,228 observations and 129 variables
```

[top](#top)

#Frequency Table of Incident Complaint
```{r}
symptoms %>%
  group_by(incident_complaint) %>%
  freq_table()
```


#Cross Tabulation of Incident Complaint by Symptoms
* I am having a hard time creating this table, because the symptoms are binary columns and not just one column.  This is the code I have thus far.


```{r}
symptoms %>%
  group_by(incident_complaint, symptoms_simplified) %>%
  freq_table() 

```









# Session information

```{r echo=FALSE}
rm(list = ls())
```

```{r echo=FALSE}
sessionInfo()
```