---
title: "Descriptive Analysis for Evaluation of the DETECT Elder Mistreatment Screening Tool"
date: "Created: 2020-03-31 <br> Updated: `r Sys.Date()`"
---

# Overview

Calculation of some descriptive statistics for the Evaluation of the DETECT Elder Mistreatment Screening Tool manuscript


# Load packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
Sys.setenv(TZ = "US/Central")
```

```{r message=FALSE}
library(dplyr)
```


# Read-in data

* Make sure to connect to the UTHealth server using the path stored in NOTES.

* aps_02_variable_management.rds was created in data_aps_02_variable_management.Rmd

```{r}
aps <- readr::read_rds("/Volumes/DETECT/one_year_data/aps_02_variable_management.rds")
```

```{r}
dim(aps) # 18,080 observations and 64 variables
```

# Dates included

Check to see the minimum and maximum dates included in this data.

```{r}
aps %>% 
  summarise(
    `Earliest date in data` = min(intake_date) %>% as.Date(),
    `Latest date in data` = max(intake_date) %>% as.Date()
  )
```

Earliest date in data: 2014-12-31	   
Latest date in data: 2018-02-28	


# Unique people

Based on the unique person id sent to me by APS

```{r}
length(unique(aps$person_id))
```


# Mean age

Get the mean age by report, not person.    
1. Our unit of analysis is the report.   
2. A person's age can vary over time in this dataset.    

```{r}
aps %>% 
  summarise(
    mean_age = mean(age),
    sd_age = sd(age)
  )
```



# Session information

```{r echo=FALSE}
sessionInfo()
```